---
title: "MSCR 534: Analysis Exercise 2"
author: "Anish Shah"
date: "April 28, 2020"
output: 
  pdf_document:
    latex_engine: xelatex
always_allow_html: yes
header-includes:
  - \usepackage{dcolumn}
  - \usepackage{float}
  - \usepackage{graphicx} 
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{caption}
  - \captionsetup[table]{labelformat=empty}
---

```{r global_options, include = FALSE}
# Knitr options
knitr::opts_chunk$set(
  cache = TRUE,
  warning = FALSE,
  eval = TRUE,
  echo = TRUE,
  include = TRUE,
  message = FALSE,
  dpi = 600,
  dev = "png",
  options("scipen" = 999, "digits" = 3),
  tinytex.verbose = TRUE,
  tidy = FALSE
)

options(xtable.comment = FALSE)

library(tidyverse)
library(knitr)
library(rmarkdown)
library(magrittr)
library(compareGroups)
library(haven)
library(kableExtra)
library(DiagrammeR)
library(stargazer)
library(DescTools)
library(data.table)
library(epitools)
library(broom)
library(sandwich)
library(lmtest)
library(Hmisc)
library(splines)
library(riskRegression)
library(prodlim)
```

# Assignment Description

1. To design and create tables for presentation of epidemiologic data.
1. To perform Poission regression, regression with splines, and use competing risk analysis with Cox Hazards Regression. 
1. To generate and interpret measures of association using SAS.

\clearpage

# Question 1: Poisson Regression

A researcher aimed to compare the automobile mortality rates among drivers aged 18-70 between two states (state 1 and state 0) during the year 2003. The total number count of deaths (see variable _deaths_) due to automobile accidents within each of eight age groups are listed in the datelines. The age groups are 18-22, 23-35, 36-41, 42-47, 48-53, 54-58, 59-63, and 64-70. The average number of people in each age group are included in the variable _population_ which was determined by the states Department of Motor Vehicles database on number of drivers licenses.

```{r read SAS, message = FALSE, include = FALSE}
# Sas datalines variable
txt <- ("
data q1;
input deaths State age1 age2 age3 age4 age5 age6 age7 population agegp;
datalines; 
12	0	1	0	0	0	0	0	0	182756	1
14	0	0	1	0	0	0	0	0	190332	2
28	0	0	0	1	0	0	0	0	201221	3
65	0	0	0	0	1	0	0	0	876231	4
102	0	0	0	0	0	1	0	0	921923	5
100	0	0	0	0	0	0	1	0	1298922	6
133	0	0	0	0	0	0	0	1	1283281	7
120	0	0	0	0	0	0	0	0	1018728	8
25	1	1	0	0	0	0	0	0	161222	1
43	1	0	1	0	0	0	0	0	159594	2
144	1	0	0	1	0	0	0	0	604213	3
323	1	0	0	0	1	0	0	0	809765	4
367	1	0	0	0	0	1	0	0	1098191	5
298	1	0	0	0	0	0	1	0	1589381	6
301	1	0	0	0	0	0	0	1	948312	7
121	1	0	0	0	0	0	0	0	728193	8
;
run;
")

# Parse it (save data as q1)
ParseSASDatalines(txt)
```

## Table 1

_Fill in Table 1. Use proc genmod to obtain the rate ratios and 95% confidence intervals._

```{r tbl 1}
# Dt 
dt <- as.data.table(q1)

# Crude table
crude <- dt[, .(dead = sum(deaths), pop = sum(population)), by = State, with = TRUE]

# Crude rates
crude$dead / (crude$pop / 100000)

# Summary table for crude rate ratios
epitab(crude$dead, crude$pop, method = "rateratio")

# Adjusted rates
dt[, rates := deaths/(population/100000)]
dt$ln_pop <- log(dt$population)
dt[, agegp := factor(agegp)]
m <- glm(deaths ~ State + agegp + offset(ln_pop), family = poisson(link = "log"), data = dt)

# Present data
tidy(m, exponentiate = TRUE, conf.int = TRUE) %>%
	kable(format = "latex", booktabs = TRUE, caption = "Adj RR") %>%
	kable_styling(latex_options = "HOLD_position")
```
	
\clearpage

# Question 2: Splines

For Question 2, use the SAS dataset _cohort_ from Lab 8. In Lab 8 we used four equal cut-points and a quadratic spline in a Cox Hazards regression using a SAS macro. For Question 2, redo this spline regression analysis by using the same SAS macro with 3 equal knots (tertiles). Use the data to complete Table 2 and answer Questions 2 below.

```{r data}
cohort <- read_sas("cohort.sas7bdat")

# Unadjusted
m <- coxph(Surv(timetooc, outc2) ~ baserbg2, data = cohort)

# Tertiles
cohort$glugrp <- cut2(cohort$baserbg2, g = 3) %>% factor()
cohort$agecat %<>% factor()
m3 <- coxph(Surv(timetooc, outc2) ~ glugrp, data = cohort)
m3adj <- coxph(Surv(timetooc, outc2) ~ glugrp + agecat, data = cohort)

# Splines
k <- quantile(cohort$baserbg2, p = c(0.25, 0.50, 0.75))
ms <- coxph(Surv(timetooc, outc2) ~ bs(baserbg2, knots = k, degree = 1), data = cohort)
coxph(Surv(timetooc, outc2) ~ bs(baserbg2, knots = k, degree = 1), data = cohort) %>% summary()
tidy(ms, exponentiate = TRUE, conf.int = TRUE)
```

\clearpage

# Question 3: Competing Risk Analysis

A cohort study aimed to determine the hazard rate of incident of invasive pulmonary Aspergillosis (fungal infection) in patients with idiopathic pulmonary fibrosis. The primary exposure of interest was biologic sex and a competing risk of concern was all-cause mortality. Assume all participants were free of Aspergillosis at baseline and were at risk of disease.

```{r question 3 data}
q3 <- read_sas("analysisexercise2question3.sas7bdat") %>%
	as.data.table()
cols = c("sex", "censor1")
q3[, (cols) := lapply(.SD, as.factor), .SDcols = cols]
```

What is the crude risk ratio for incident Aspergillosis, comparing men to women?

```{r}
# Crude risk
m <- glm(censor1 ~ sex, family=binomial("log"), q3[censor1 != 2])
tidy(m, exponentiate = TRUE, conf.int = TRUE) %>%
	kable(format = "latex", booktabs = TRUE, caption = "Crude RR") %>%
	kable_styling(latex_options = "HOLD_position")

# Median time
tmp <- median(q3$timetoc2)
```

## Table 3A

```{r without competing risks adjusted for}
# Turn death into potential event (dropping competing risks)
q3$censor <- ifelse(q3$censor1 == 1, 1, ifelse(q3$censor1 == 2, 1, 0))

# Crude HR
m <- coxph(Surv(timetoc2, censor) ~ sex, data = q3)
tidy(m, exponentiate = TRUE, conf.int = TRUE) %>%
	kable(format = "latex", booktabs = TRUE, caption = "Crude HR") %>%
	kable_styling(latex_options = "HOLD_position")

# Adjusted
m <- coxph(Surv(timetoc2, censor) ~ sex + age + smk, data = q3)
tidy(m, exponentiate = TRUE, conf.int = TRUE)  %>%
	kable(format = "latex", booktabs = TRUE, caption = "Crude HR") %>%
	kable_styling(latex_options = "HOLD_position")
```

## Table 3B

```{r with competing risks}
### Cause specific for aspy
q3$censor <- ifelse(q3$censor1 == 1, 1, 0)

# Crude HR
m <- coxph(Surv(timetoc2, censor) ~ sex, data = q3)
tidy(m, exponentiate = TRUE, conf.int = TRUE) %>%
	kable(format = "latex", booktabs = TRUE, caption = "Crude Cause = Asp") %>%
	kable_styling(latex_options = "HOLD_position")

# Adjusted
m <- coxph(Surv(timetoc2, censor) ~ sex + age + smk, data = q3)
tidy(m, exponentiate = TRUE, conf.int = TRUE) %>%
	kable(format = "latex", booktabs = TRUE, caption = "Adj Cause = Asp") %>%
	kable_styling(latex_options = "HOLD_position")

### Cause specific for dead
q3$censor <- ifelse(q3$censor1 == 2, 1, 0)

# Crude HR
m <- coxph(Surv(timetoc2, censor) ~ sex, data = q3)
tidy(m, exponentiate = TRUE, conf.int = TRUE) %>%
	kable(format = "latex", booktabs = TRUE, caption = "Crude Cause = Dead") %>%
	kable_styling(latex_options = "HOLD_position")

# Adjusted
m <- coxph(Surv(timetoc2, censor) ~ sex + age + smk, data = q3)
tidy(m, exponentiate = TRUE, conf.int = TRUE) %>%
	kable(format = "latex", booktabs = TRUE, caption = "Adj Cause = Dead") %>%
	kable_styling(latex_options = "HOLD_position")

### Subdistribution for asp

# Crude
m <- FGR(Hist(timetoc2, censor1, cens.code = 0) ~ sex, cause = 1, data = q3)

# Adjustd
m <- FGR(Hist(timetoc2, censor1, cens.code = 0) ~ sex + age + smk, cause = 1, data = q3)

### Subdistribution for death

# Crude
m <- FGR(Hist(timetoc2, censor1, cens.code = 0) ~ sex, cause = 2, data = q3)

# Adjustd
m <- FGR(Hist(timetoc2, censor1, cens.code = 0) ~ sex + age + smk, cause = 2, data = q3)

```

For the subdistribution model, with outcome of death, here are the summary of failure outcomes.

```{r}
x <- addmargins(table(q3$censor1))
names(x) <- c("Censored", "Competing Event", "Event of Interest", "Total")
x %>% 
	kable(caption = "Summary: Failure Outcomes", format = "latex") %>%
	kable_styling(latex_options = "HOLD_position")
```

\clearpage
