---
title: "MSCR 534: Final Exam"
author: "Anish Shah"
date: "May 1, 2020"
output: 
  pdf_document:
    latex_engine: xelatex
always_allow_html: yes
header-includes:
  - \usepackage{dcolumn}
  - \usepackage{float}
  - \usepackage{graphicx} 
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{caption}
  - \captionsetup[table]{labelformat=empty}
---

```{r global_options, include = FALSE}
# Knitr options
knitr::opts_chunk$set(
  cache = TRUE,
  warning = FALSE,
  eval = TRUE,
  echo = TRUE,
  include = TRUE,
  message = FALSE,
  dpi = 600,
  dev = "png",
  options("scipen" = 999, "digits" = 3),
  tinytex.verbose = TRUE,
  tidy = FALSE
)

options(xtable.comment = FALSE)

library(tidyverse)
library(knitr)
library(rmarkdown)
library(magrittr)
library(compareGroups)
library(haven)
library(kableExtra)
library(DiagrammeR)
library(stargazer)
library(DescTools)
library(data.table)
library(epitools)
library(broom)
library(sandwich)
library(lmtest)
library(Hmisc)
library(splines)
library(riskRegression)
library(prodlim)
library(janitor)
library(epiR)
library(mediation)
library(survival)
library(survminer)
library(viridis)
```

# Question 1

Answered in word document.

# Question 2

```{r}
B0 = -4.4195 # Intercept
B1 = 1.0221 # FPM exposure
B2 = 0.4210 # age <60
B3 = 0.5600 # age >= 60
B4 = 0.7712 # FPM with young age
B5 = 0.8211 # FPM with old age

# 65 year old with high FPM exposure
estimate = B0 + B1 + B3 + B5
cumInc = exp(estimate)

# 38 year old with low FPM exposure
estimate = B0 + B2
cumInc = exp(estimate)

# 42 with High FPM compared to 70 with low FPM
RR = exp(B1 + B2 + B4) / exp(B3)
```

# Question 3

Answered in word document.

# Question 4

Answered in word document.

# Question 5

Table 3 filled out using this information.

```{r describe the hip fx}
# Data
q5 <- read_sas("q5.sas7bdat") %>% 
	clean_names() 

df <- q5

# Risk in hip fracture
df$hip55 %<>% factor(levels = c(1,0), labels = c("Fx", "NoFx"))
df$livea %<>% factor(levels = c(1,0), labels = c("Alone", "Others"))

tbl <- table(df$livea, df$hip55)
epitbl <- epi.2by2(tbl)

# Print
print(epitbl)
```

Verified odds ratio via comparison of groups.

```{r compare the hip fx, results='asis'}
# Crude risk of hip fracture
compareGroups(hip55 ~ livea, data = df) %>%
	createTable(show.all = TRUE, show.ratio = TRUE) %>%
	export2md(format = "latex", caption = "Comparison of Hip Fx and Loneliness")
```

Adjusted OR done through logistic regression.

```{r regress the hip fx, results='asis'}
m <- glm(hip55 ~ livea + alc + ager, family = binomial("logit"), data = q5)

# Make pretty
stargazer(m, header = FALSE, type = "latex", 
					apply.coef = exp, 
					p.auto = FALSE,
					report = "vc*s", title = "Adjusted OR for Hip Fx",
					ci.custom = list(exp(confint(m))))

```

# Question 6

We will need to establish several relationships to study the mediation effects of depression on the relationship between living alone and hip fractures. 

We can test the relationship of living alone with depression. 

```{r sad and lonely}
# Data
df <- q5

# Tidy
df$livea %<>% factor(levels = c(0,1), labels = c("Others", "Alone"))
df$dep %<>% factor(levels = c(0,1), labels = c("Happy", "Sad"))
df$hip55 %<>% factor(levels = c(0,1), labels = c("NoFx", "Fx"))

compareGroups(dep ~ livea, data = df) %>%
	createTable(show.ratio = TRUE, show.all = TRUE) %>%
	export2md(format = "latex", caption = "Loneliness and Depression")
```

We can then test the relationship of the potential mediator of depression on hip fractures.

```{r}
# Data as above
compareGroups(hip55 ~ dep, data = df) %>%
	createTable(show.ratio = TRUE, show.all = TRUE) %>%
	export2md(format = "latex", caption = "Depression and Fx")
```

Now, since depression is associated with both exposure and outcome, we can test it as a mediator using an adjusted regression model.

```{r, results='asis'}
# Three models
m1 <- glm(hip55 ~ livea, family = binomial("logit"), data = q5)
m2 <- glm(hip55 ~ dep, family = binomial("logit"), data = q5)
m3 <- glm(hip55 ~ dep + livea, family = binomial("logit"), data = q5)

# Make pretty
stargazer(m1, m2, m3, header = FALSE, type = "latex", 
					apply.coef = exp, 
					p.auto = FALSE,
					report = "vc*s", title = "Mediation Analysis of Loneliness, Sadness, and Fractures",
					ci.custom = list(exp(confint(m1)), exp(confint(m2)), exp(confint(m3)))
					)
```

To mesure effects, we need to look at measured relationships in the causal model.

- Living Alone -> Depression
- Living Alone + Depression -> Hip Fx

```{r}
# Models
m <- glm(dep ~ livea, family = binomial("logit"), data = q5)
y <- glm(hip55 ~ livea + dep, family = binomial("logit"), data = q5)

# Mediation
res <- mediate(m, y, treat = "livea", mediator = "dep", boot = TRUE, sims = 500)
summary(res)
```

This supports the idea the depression only accounts for a very small part (indirect effect) of the total effect of living alone on hip fractures.

# Question 7

This creates a Kaplan Meier curve for assessing lung cancer incidence by smoking strata.

```{r take in q7 data, fig.align='center'}
q7 <- read_sas("q7.sas7bdat") %>%
	clean_names()

# Create a KM graph (exclude death during f/u)
df <- subset(q7, outcome != 1)
df$outcome[df$outcome == 2] <- 1
df$smoke %<>% factor()
fit <- survfit(Surv(survival_t, outcome) ~ smoke, data = df)
ggsurvplot(
	fit, 
	data = df, 
	risk.table = TRUE,
	pval = TRUE,
	conf.int = TRUE,
	ggtheme = theme_minimal(),
	palette = viridis_pal(option = "E")(2)
)
```

Now we need to calculate Cox regression to create a hazard rate ratio, as well the subdistribution models.

```{r}
# Data excluding those who died from analysis, as above
df <- subset(q7, outcome != 1)
df$outcome[df$outcome == 2] <- 1
df$smoke %<>% factor()
df$agecat %<>% factor()
df$sex %<>% factor()

# Cox regression
m1 <- coxph(Surv(survival_t, outcome) ~ smoke + agecat + sex, data = df)
tidy(m1, exponentiate = TRUE, conf.int = TRUE) %>%
	kable(format = "latex", 
				booktabs = TRUE, 
				caption = "Adjusted HR for Lung Cancer") %>%
	kable_styling(latex_options = "HOLD_position")

# Subdistribution for lung cancer
df <- q7
df$smoke %<>% factor()
df$agecat %<>% factor()
df$sex %<>% factor()

m2 <- FGR(Hist(survival_t, outcome, cens.code = 0) ~ smoke + agecat + sex, 
					cause = 2, 
					data = df)
print(m2)

# Subdistribution for death
m3 <- FGR(Hist(survival_t, outcome, cens.code = 0) ~ smoke + agecat + sex, 
					cause = 1, 
					data = df)
print(m3)
```

