---
title: "MSCR 509: High Dimensional Analysis"
subtitle: "Homework 11"
author: Anish Shah
date: April 20, 2020
output: pdf_document
latex_engine: xelatex
header-includes:
  - \usepackage{dcolumn}
  - \usepackage{float}
  - \usepackage{graphicx} 
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
always_allow_html: true
---

```{r setup, global_options, include=FALSE}
knitr::opts_chunk$set(
	cache = TRUE,
	warning = FALSE,
	eval = TRUE,
	echo = FALSE,
	include = FALSE,
	message = FALSE,
	dpi = 600,
	dev = "png",
	options("scipen" = 999, "digits" = 4)
)

library(knitr)
library(tinytex)
library(rmarkdown)
library(tidyverse)
library(broom)
library(haven)
library(magrittr)
library(lmtest)
library(compareGroups)
library(pROC)
library(stargazer)
library(caret)
library(e1071)
library(kableExtra)
library(ResourceSelection)
library(glmulti)
library(cvAUC)
library(glmnet)
library(MASS)
library(fdrtool)
library(qvalue)
library(mice)
library(factoextra)
library(data.table)
library(Hmisc)
```

# Question 1

The file places.txt contains data from Places Rated Almanac data (Boyer and Savageau) which rates 329 metropolitan areas of the United States according to nine composite variables:

- Climate and Terrain 
- Housing 
- Health Care & Environment 
- Crime 
- Transportation 
- Education 
- The Arts 
- Recreation 
- Economics 

Each composite variable is constructed from available data, for example, housing costs are the sum of 3 components: utility bills, property taxes, and mortgage payments. Utility bills are a function of gas and electric prices, heating degree days (when the temperature is less than 65 0 F), air conditioning degree days (when the temperature is more than 800 ), and how common it is for houses to be heated electrically rather than by gas.  

Notes:

- The data for many of the variables are strongly skewed to the right. 
- The natural log transformation was used to normalize the data.
- The places_data.sas file has code to read dataset and log transform 9 variables.

Researchers are interested in learning ranking of places to live in the United States.  One way of doing this is to principle component analysis.

```{r}
# Read in places.txt file
places <- fread("places.txt")

# Name columns
columns <- Cs(climate, housing, health, crime, trans, educate, arts, recreate, econ, id)
names(places) <- columns

# Log transform data
places[, (columns) := lapply(.SD, log), .SD = columns]
```

## Part 1

_Report the principal components (i.e., the loadings associated with each PC) of these variables based on correlation matrix (i.e. without COV option)._

```{r, include = TRUE}
# Principal components and loadings based on cor matrix
# Eigenvalues/vectors
pca <- prcomp(places, scale = TRUE, center = TRUE)

# print loadings
kable(pca$rotation, format = "latex", caption = "Loadings of PC") %>%
	kable_styling(font_size = 8)
```

## Part 2

_How many principal components would you select to summarize this data? Explain why?_

```{r, include = TRUE}
fviz_screeplot(pca)
```

Based on the scree plot of the variance, we can see that by the __PC4__ we have dipped below the eigenvalue of 1, thus limiting the amount of novel information we are receiving. The first three PC would explain approximately 56% of the total variance, thus may be sufficient in our analysis. 


## Part 3

_What are the interpretations of the first three PCs?_

When we look at the contributing loadings of PC1-3, we see that PC1 and PC2 use a linear combination of all the variables for the most part. However, PC3 only relies on several variables: _climate, trans, educate, econ, and id_. We can however summarize and say that hte first 3 PC explain almost 6% of the variance, and thus may be a sufficient to start modeling to predict rankings of places to live.

## Part 4

_Explain why PCA is a good technique for this problem._

PCA is a good technique for this problem because there are 10 attributes that are contributing to rankings, however each of them are likely to have some level of correlation. A regression model for rankings will likely be biased and underpowered at estimating the best rankings because of overfitting (with too many variables). With just PC1-3 we can make a fairly well fitting model.
